Drivers of tree regeneration remain challenging to be captured by models of

### 2022-11-23 manuscript/manuscriptRW.Rmd
mainly more changes in the manuscript  discussion section

---
title: "Drivers of tree regeneration remain challenging to be captured by models of forest dynamics"
author: 
- Olalla Díaz-Yáñez, 
- Yannek Käber, 
- Mats Mahnken, 
- Samuel Fischer, 
- Xavier Morin, 
- Manfred Lexer, 
- Christian Hochauer, 
- Katarina Merganicova, 
- Tobias Mette, 
- Werner Rammer, 
- Kristin H. Braziunas, 
- Paola Mairota, 
- Josef Bruna,
- Heike Lischke, 
- Daniel Scherrer, 
- Simon Scheiter, 
- Thomas Hickler, 
- Tim Anders,
- Jessica Hetzer, 
- Harald Bugmann
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
  word_document:
    reference_docx: template_paper.docx
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: draft.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

@@

Regeneration is a key process in forest dynamics. Models of forest
dynamics have been long lagging an assessment in their capacities to
properly represent this process. Using 15 models and variants of forest
dynamics at the stand, landscape and global levels, we simulate the
regeneration across 200 sites representing environmental gradients
across Europe. The simulation results were compared with observed data
collected from forest reserves across Europe. In this study we want to
asses how well are models of forest dynamics capturing regeneration
levels, species composition and mortality in regeneration. The results
show that most of the models of forest dynamics tend to over-estimate
recruitment rates although this is not always connected to large
mortality rates in early dynamics stages. The effect of light
availability on recruitment rates is better captured than temperature or
soil moisture, however the patterns are not consistent across models.
Species richness matches quite well compared to the ranges observed in
the data. Models simulating higher species diversity at the stand level
do not feature higher diversity in recruitment. *Abies alba* was well
captured across temperature and soil moisture gradients by most models
and *Fagus sylvatica* was generally underestimated. *Picea abies*
recruitment was generally over-estimated on colder sites. HERE TWO
SENCENTECES DISCUSSION/CONCLUSION. HERE A SO WHAT SENTENCE

@@

Tree recruitment is a fundamental process in forest dynamics, in
particular with increasing disturbance under climate change. Correctly
capturing regeneration in forest dynamics is fundamental to for example
evaluate post-disturbance dynamics as this will define the status of the
forests in the first 15-20 years [@seidl_post-disturbance_2022]. If this
initial state it is not well captured then we can not properly assess
aspects such as the forest resilience, or changes in carbon capture. In
this study, for the first time, projections of tree recruitment done by
several models of forest dynamics were confronted with observed data
across a large environmental gradients. The observed data that we used
[@käber2022] is exceptional in this regard, and to use it for model
benchmarking has gave us much more insights than if we have used it for
calibration.

Adequate tree recruitment rates define the structure and composition of
forests. Most of the models of forest dynamics tended to over-recruit
trees in all the sites. Over estimating recruitment might have
fundamental implications on the role of regrowth and forest carbon, with
an overestimation in their capacity for carbon capture for example after
the impact of disturbances [@pugh2019]. Excess in recruitment also
requires of larger mortality rates in later stages, changing the
expected forest densities at early stages and forcing the models to have
stronger filters. Regeneration levels define the success of future
forests, but it is equally important to correctly identify un-stocked
forest areas [@rammer2021]. Tree recruitment is often characterized by
an excess number of zero counts [@fortin2007]. In this study some of the
models did not have lack of recruitment or their proportion was very
low. This substantial difference with the observed data might be a
consequence of assuming that the observed data is also in equilibrium,
where lack of recruitment in sites could also be an indication of
"non-equilibrium" due to management history.

Light availability effect on tree regeneration have been widely
documented [@messier1999; @collet2006; @berdanier2016]. However it is
difficult to have variables defining light availability for the
regeneration. In this study, we used the total stand basal area as a
proxy for light availability. There was differences in the basal area
ranges, some of the models had stand basal areas lower than the observed
data but there was also models that did have more stocked stands than
the observed data, therefore it was not possible to evaluate the
recruitment for lower density stands. Regeneration at low stand
densities is an interesting aspect to evaluate how well forests are
recovering e.g. after the creation of gaps due to disturbances. We would
have expected that the recruitment of the different species would become
sparser and droped out completely as a function of a proxy of light
availability, however few models showed this trend indicating that the
relation between recruitment and light availability it is not yet
correctly defined.

When tree composition in recruitment is not well captured, then we can
not assess the functional composition of the forest appropriately, e.g.
their sensitive to drought or resilience to new disturbances events
[@redmond2015; @seidl_post-disturbance_2022]. In the models simulation
outputs overall species richness levels in recruitment were inside the
observed ranges, however there were variations on how the main tree
species were represented in the environmental gradient. Most models
maintain or even increase in recruitment diversity between the 7 and 10
cm thresholds, and the same is visible from the observed data, however
the differences are not significant. Species richness variations might
be explained by the way each model considers recruitment, e.g. in the
absence of adults, recruitment can be linked to either seed
rain-dispersal, or to seed/seedling bank, but also on what processes are
formulated with stochastic or deterministic approach. Differences might
also be expected between cohort and individual trees based models
**how?**. However, our results do not show differences between models
with an without feedback in recruitment, **why?** The adult forests
species richness was better captured than the recruitment species
richness, and only few models over-predicted stand level species
richness **why these ones?**.

We could expect that models that over-recruit trees will also have a
large mortality between 7 and 10 cm thresholds but this was not shown.
This implies that these models must have the excess of mortality in
later stages. Higher mortality in the adult component might enhance the
share of less shade tolerant species for those models. There might be a
difference between models including stochastic processes in recruitment
and those which do not ??. There was also a group of models below the
Reineke line indicating a lower mortality than expected by the
self-thinning rule, which may point to facilitation, or simply a higher
mortality rate before the trees reached 7 cm diameter, another
possibility would have been that these models underestimate recruitment
which was not the case.

There is also the question of how much level of detail one have to give
to define regeneration processes [@konig_tree_2022; @bugmann2022].
Models with more detailed and complex recruitment modules, will have
also increase parametrisation which will lead to lower generalization as
this values will have to be collected from specific geographical
regions. This is significant relevant when looking at mixed forests and
the species level parametrization. **How it relates to the results?
Future model development recommendations?**

-   Difference in observations number between observed data (2-4 samples
    per site) and the simulated data (200 samples per site) and the
    implication on variability.

-   Assumption equilibrium it is expected overestimation because we do
    not know the status of the observed data more certainly not in
    equilibrium etc

-   methodological limitations compare mortality with the self thinning
    line etc Values below the 1.0 line show sites where there is a
    higher number of 10 cm trees than in 7 cm - this is an artifact of
    the method - we expect space for time substitution, but in reality
    we just have one measurement.

-   choice of an ingrowth limit of 10 /7cm DBH. if the model gets
    overall species composition right, it must also get ingrowth of
    these species rather large trees right. a better test would be for
    the ingrowth limit to be lower. we chose the highest calipering
    limit of 10 to maximize data availability

-   methodology number of samples in observations vs simulations per
    site

-   We use all the models to rpoduce values per 1ha, stand ang global
    models potentially disregarding the advantages that landscape models
    present with spatial seed dispersal (TreMig, iLand and LAndClim
    references). Without spatial context we are limited to the seed
    influx from the stand itself if the model does not have a background
    seed input. But even global vegetation models are lacking this
    dispersal between cells [@hanburybrown_forest_2022] abstraction of
    space or even the trees themselves. It is important to include
    spatial considerations on tree recruitmen [@beckage2003]

-   Comparing PFTs with species. We lose the species specific
    information on with substantial drawbacks [@purves2008]

-   The loss of species-specific environmental responses through the
    collation of species communities into PFTs has been widely accepted,
    as parameterization efforts would otherwise exceed the available
    means. However, it is doubtful whether the definition of PFTs
    provides sufficient flexibility for an adequate representation of
    the ecological processes and the differences between species

-   Recruitment of trees into size classes of larger 4 cm DBH, as in our
    four data sets, represents a relatively late stage of forest
    regeneration. It integrates over sometimes long periods since
    germination and multiple environmental constraints act upon young
    trees of widely different size. kaber 2021

-   We do not know what happen before 7cm in the observed data and does
    not make sense to look at that in the simulated data. 7cm is a large
    sapling, consequences of not having that information to asses forest
    dynamics. Recommendations on data collection.

@@

Capture properly recruitment rates in the environmental gradient might
impact future forests.

>>>>

---
title: Drivers of tree regeneration remain challenging to be captured by models of
  forest dynamics
author:
- Olalla Díaz-Yáñez,
- Yannek Käber,
- Mats Mahnken,
- Samuel Fischer,
- Xavier Morin,
- Manfred Lexer,
- Christian Hochauer,
- Katarina Merganicova,
- Tobias Mette,
- Werner Rammer,
- Kristin H. Braziunas,
- Paola Mairota,
- Josef Bruna,
- Heike Lischke,
- Daniel Scherrer,
- Simon Scheiter,
- Thomas Hickler,
- Tim Anders,
- Jessica Hetzer,
- Harald Bugmann
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
  word_document:
    reference_docx: template_paper.docx
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: draft.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

@@

Regeneration is a key process in forest dynamics. Models of forest
dynamics have been long lagging an assessment in their capacities to
properly represent this process. Using 15 models and variants of forest
dynamics at the stand, landscape and global levels, we simulate the
regeneration across 200 sites representing environmental gradients
across Europe. The simulation results are compared with observed data
collected from forest reserves across Europe. In this study we want to
asses how well are models of forest dynamics capturing regeneration
levels, species composition and mortality in tree recruitment. The
results show that most of the models of forest dynamics tend to
over-estimate recruitment rates although this is not always connected to
large mortality rates in early dynamics stages. The effect of light
availability on recruitment rates is better captured than temperature or
soil moisture, however the patterns are not consistent across models.
Species richness matches quite well compared to the ranges observed in
the data. Models simulating higher species diversity at the stand level
do not feature higher diversity in recruitment. *Abies alba* was well
captured across temperature and soil moisture gradients by most models
and *Fagus sylvatica* was generally underestimated. *Picea abies*
recruitment was generally over-estimated on colder sites. HERE TWO
SENCENTECES DISCUSSION/CONCLUSION. HERE A SO WHAT SENTENCE

@@

Tree recruitment is a fundamental process in forest dynamics, in
particular with increasing disturbance under climate change. Correctly
capturing regeneration in forest dynamics is fundamental to for example
evaluate post-disturbance dynamics as this will define the status of the
forests in the first 15-20 years [@seidl_post-disturbance_2022]. If this
initial state it is not well captured then we can not properly assess
aspects such as the forest resilience, or changes in carbon capture. In
this study, for the first time, projections of tree recruitment done by
several models of forest dynamics were confronted with observed data
across a large environmental gradients.

The observed data that we used [@käber2022] is exceptional in this
regard, and to use it for model benchmarking has gave us insights on
recruitment levels, species composition and mortality and overall a
broader understanding than if we have used these data for a model
calibration. The lower available threshold in the observed data of 7 cm
provides a limit from which we can understand tree recruitment, which
might be a rather long period since germination as relevant
environmental constraint might have already act on those young trees
[@käber2021]. However recruitment data with lower thresholds are not
available in a harmonized manner from large environmental gradients. The
observed data was collected from forest reserves where there has not
been management for long periods of time, this might help to properly
assume that these structures are in an equilibrium state. This
assumption might also justify some of the rather large recruitment rates
that we are seeing in the models projections compare to the observed
ones.

Adequate tree recruitment rates define the structure and composition of
forests. Most of the models of forest dynamics tended to over-recruit
trees in all the sites. Over estimating recruitment might have
fundamental implications on the role of regrowth and forest carbon, with
an overestimation in their capacity for carbon capture for example after
the impact of disturbances [@pugh2019]. Excess in recruitment also
requires of larger mortality rates in later stages, changing the
expected forest densities at early stages and forcing the models to have
stronger filters. Regeneration levels define the success of future
forests, but it is equally important to correctly identify un-stocked
forest areas [@rammer2021]. Tree recruitment is often characterized by
an excess number of zero counts [@fortin2007]. In this study some of the
models did not have lack of recruitment or their proportion was very
low. This substantial difference with the observed data might also be a
consequence of assuming that the observed data is also in equilibrium,
where lack of recruitment in sites could also be an indication of
"non-equilibrium" due to management history.

Light availability effect on tree regeneration have been widely
documented [@messier1999; @collet2006; @berdanier2016]. However it is
difficult to collect variables defining light availability on
regeneration. In this study, we used the total stand basal area as a
proxy for light availability. There was differences in the basal area
ranges, some of the models had stand basal areas lower than the observed
data but there was also models that did have more stocked stands than
the observed data, therefore it was not possible to evaluate the
recruitment for the extremes of the density ranges in some models.
Regeneration at low stand densities is interesting to assess how well
forests are recovering e.g. after the creation of gaps due to
disturbances. We would have expected that the recruitment of the
different species would become sparser and droped out completely as a
function of a proxy of light availability, however few models showed
this trend indicating that the relation between recruitment and light
availability it is not yet correctly captured.

Correctly capturing tree composition in recruitment help us to assess
the functional composition of the forest appropriately, e.g. their
sensitive to drought or resilience to new disturbances events
[@redmond2015; @seidl_post-disturbance_2022]. In the models simulation
outputs overall species richness levels in recruitment was
underestimated but inside the observed ranges, however there were
variations on how the main tree species were represented in the
environmental gradient. The lower species richness in recruitment might
be a consequence of models running without disturbances, disturbances
would typically allowed for larger gaps sizes allowing less shade
tolerant species to grow in larger proportions. Most models maintain or
even increase in recruitment diversity between the 7 and 10 cm
thresholds, and the same is visible from the observed data, however the
differences are not significant. Species richness variations might be
explained by the way each model considers recruitment, e.g. in the
absence of adults, recruitment can be linked to either seed
rain-dispersal, or to seed/seedling bank, but also on what processes are
formulated with stochastic or deterministic approach. Differences might
also be expected between cohort and individual trees based models
**how?**. However, our results do not show differences between models
with and without feedback in recruitment, **why?** The adult forests
species richness was better captured than the recruitment species
richness, and only few models over-predicted stand level species
richness **why these ones?**.

We could expect that models that over-recruit trees will also have a
large mortality between 7 and 10 cm thresholds but this was not shown.
This implies that these models must have the excess of mortality in
later stages. Higher mortality in the adult component might enhance the
share of less shade tolerant species for those models. There might be a
difference between models including stochastic processes in recruitment
and those which do not **??**. There was also a group of models below
the Reineke line indicating a lower mortality than expected by the
self-thinning rule, which may point to facilitation, or simply a higher
mortality rate before the trees reached 7 cm diameter, another
possibility would have been that these models underestimate recruitment
which was not the case. This might also be partly an artifact of the
method, as we are expecting space for time substitution, but in reality
we just have one measurement.

There is also the question of how much level of detail one have to give
to define regeneration processes [@konig_tree_2022; @bugmann2022].
Models with more detailed and complex recruitment modules, will have
also increase parametrisation which will lead to lower generalization as
this values will have to be collected from specific geographical
regions. This is significant relevant when looking at mixed forests and
the species level parametrization. **How this relates to the results?
Future model development recommendations?**

In these study we considered a wide range of models of forest dynamics,
from stand to global scale. However, the simulation set up was limited
to 1 ha recruitment values. It is important to include spatial
considerations on tree recruitment [@beckage2003] and the lack of it
might represent a more important disadvantage for landscape level
models. Without spatial context we are limiting recruitment to the seed
influx from the stand itself if the model does not have a background
seed input. But even global vegetation models are lacking this dispersal
between cells where there is an abstraction of the space or even the
trees themselves [@hanburybrown_forest_2022]. The selected models also
had limitations on species representation, some models where not able to
simulate certain species due to lack of species parametrization. We also
had one model that simulate PFTs providing a less flexible comparison
the with species values; with the lose of the species specific
information might also represent substantial ecological
representativeness drawbacks [@purves2008].

@@

Capture properly recruitment rates, species composition and mortality in
the environmental gradient might impact future forests...
